---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hacksaw

<!-- badges: start -->
![](https://camo.githubusercontent.com/ea6e0ff99602c3563e3dd684abf60b30edceaeef/68747470733a2f2f696d672e736869656c64732e696f2f62616467652f6c6966656379636c652d6578706572696d656e74616c2d6f72616e67652e737667)
![CRAN log](http://www.r-pkg.org/badges/version/hacksaw)
[![Travis build status](https://travis-ci.org/daranzolin/hacksaw.svg?branch=master)](https://travis-ci.com/daranzolin/hacksaw)
<!-- badges: end -->

hacksaw is as an adhesive between various dplyr and purrr operations, with some extra tidyverse-like functionality (e.g. keeping NAs, shifting row values, casting, etc.)

## Installation

You can install the released version of hacksaw from GitHub with:

``` r
remotes::install_github("daranzolin/hacksaw")
```

## Split operations

### filter

```{r warning=FALSE, message=FALSE}
library(hacksaw)
library(dplyr)

iris %>% 
  filter_split(
    large_petals = Petal.Length > 5.1,
    large_sepals = Sepal.Length > 6.4) %>% 
  str()
```

### select

```{r}
iris %>% 
  select_split(starts_with("Sepal"),
               starts_with("Petal")) %>% 
  str()
```

### distinct

```{r}
starwars %>% 
  distinct_split(skin_color, eye_color, homeworld) %>% 
  str()
```

### mutate

```{r}
iris %>% 
  mutate_split(Sepal.Length2 = Sepal.Length * 2,
               Sepal.Length3 = Sepal.Length * 3) %>% 
  str()
```

### transmute 

```{r}
iris %>% 
  transmute_split(Sepal.Length * 2, Petal.Width + 5) %>% 
  str()
```


### slice

```{r}
iris %>% 
  slice_split(1:10, 11:15, 30:50) %>% 
  str()
```

Use the `var_max` and `var_min` helpers to easily get minimum and maximum values of a variable:

```{r}
iris %>% 
  slice_split(
    largest_sepals = var_max(Sepal.Length, 4),
    smallest_sepals = var_min(Sepal.Length, 4)
  )
```


## Casting

Tired of `mutate(... = as....(...))`? 

```{r}
starwars %>% cast_character(height, mass) %>% str(max.level = 2) 
iris %>% cast_character(contains(".")) %>% str(max.level = 1)
```

hacksaw also includes `cast_numeric` and `cast_logical`.

## Keeping NAs

The reverse of `tidyr::drop_na`.

```{r}
df <- tibble(x = c(1, 2, NA, NA, NA), y = c("a", NA, "b", NA, NA))
df %>% keep_na()
df %>% keep_na(x)
df %>% keep_na(x, y)
```

## Shifting row values

Shift values across rows in either direction

```{r}
df <- data.frame(
  s = c(NA, 1, NA, NA),
  t = c(NA, NA, 1, NA),
  u = c(NA, NA, 2, 5),
  v = c(5, 1, 9, 2),
  x = c(1, 5, 6, 7),
  y = c(NA, NA, 8, NA),
  z = 1:4
)
df

shift_row_values(df)
shift_row_values(df, at = 1:3)
shift_row_values(df, at = 1:2, .dir = "right")

```

