---
title: "Splitting a Data Frame with hacksaw"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Splitting a Data Frame with hacksaw}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Splitting a data frame with hacksaw is a little unusual. In brief, each expression with the `*_split` functions (e.g. `filter_split`, `select_split`, `slice_split`, etc.) gets evaluated separately and the outputs are appended to a list. This is useful when creating multiple objects from a single data frame and allows a greater level of control when splitting. Thus, the [zeallot](https://github.com/r-lib/zeallot) and [purrr](https://github.com/tidyverse/purrr) packages are great compliments to hacksaw.

For example, you can separate two separate tables from the iris dataset, one with large petals, and another with large sepals:

```{r setup}
library(hacksaw)
library(dplyr)

iris %>% 
  filter_split(
    large_petals = Petal.Length > 5.1,
    large_sepals = Sepal.Length > 6.4
    ) %>% 
  str()
```

Adding a third expression would append another table to the list:

```{r}
iris %>% 
  filter_split(
    large_petals = Petal.Length > 5.1,
    large_sepals = Sepal.Length > 6.4,
    large_everything = Petal.Length > 5.1 & Sepal.Length > 6.4
    ) %>% 
  str()
```

`select_split` functions in a similar fashion, and can leverage dply's select helpers:

```{r}
iris %>% 
  select_split(
    starts_with("Sepal"),
    starts_with("Petal"),
    everything(),
    ) %>% 
  str()
```

`mutate_split` can be useful when you want to easily adjust the function inputs from a single data frame, or when a pivoting operation eludes you (although splitting with hacksaw will almost always be a less efficient operation, as it creates multiple copies of your data frame).

```{r fig.width=7, fig.height=5}
library(ggplot2)
library(purrr)
library(tidyr)
library(patchwork)

plot_fun <- function(data, x, y, xlab) ggplot(data, aes({{x}}, {{y}})) + geom_point() + xlab(xlab)
p1 <- mtcars %>%
  mutate_split(x = cyl, x = hp, x = wt) %>%
  map2(c("cyl", "hp", "wt"), ~plot_fun(.x, x, mpg, .y)) %>%
  reduce(`+`)

p2 <- mtcars %>%
  select(mpg, cyl, hp, wt) %>%
  pivot_longer(names_to = "var", values_to = "val", cyl:wt) %>%
  ggplot(aes(val, mpg)) +
  geom_point() +
  facet_wrap(~var, nrow = 1, scales = "free")

p1 <- p1 + plot_annotation(title = "{hacksaw} + {patchwork}")
p2 <- p2 + ggtitle("pivoting + faceting")
p1/p2 + plot_annotation(title = "{hacksaw} + {patchwork}")
```

Finding the number of unique values across multiple columns is more intuitive to type to the regular tidyverse style:

```{r}
starwars %>% summarize_at(vars(name, hair_color, homeworld), ~length(unique(.)))

starwars %>% distinct_split(name, hair_color, homeworld) %>% lengths()
```

`transmute_split` may save you a few `pull`s, and `slice_split` splits your data frame by indices:

```{r}
iris %>% 
  slice_split(1:10, 11:15, 30:50) %>% 
  str()
```

Use the `var_max` and `var_min` helpers to easily get minimum and maximum values:

```{r}
iris %>% 
  slice_split(
    largest_sepals = var_max(Sepal.Length, 5),
    smallest_sepals = var_min(Sepal.Length, 5)
  )
```

